---
title: "ICMPC17 Analysis"
author: "Kyle M. Lang & Eriko Fukuda"
date: "`r Sys.time() |> format('%Y-%m-%d')`"
output: 
  prettydoc::html_pretty:
    toc: true
    df_print: paged 
    theme: architect
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center", 
                      comment = NA)

options(width = 82)
```

---

# Setup & Data Preperation

---

First, we'll prep the environment.

```{r prep}
rm(list = ls(all = TRUE))

library(magrittr)
library(dplyr)
library(lme4)
library(performance)
library(ggplot2)

source("helper_functions.R")

dataDir <- "../data/"
fn1     <- "PerMusA_N73.csv"
fn2     <- "Combined_sample size 73.csv"
```

Now, we'll load and inspect the data.

```{r data_ingest}
## Read the data: 
dat0 <- read.csv2(paste0(dataDir, fn1))

## Temporarily exclude ID (for privacy purposes):
tmp <- dat0 %>% select(-multichar_response)

## Check data:
head(tmp)
summary(tmp)
str(tmp)

## Any missing data?
colMeans(is.na(dat0))

## Frequencies of categorical variables:
table(dat0$correct)
table(dat0$response)
table(dat0$title)
```

We don't need the bird sound condition, so we'll drop those cases.

```{r drop_birds}
## Drop the bird sound condition:
dat1 <- dat0 %>%
    filter(title != "2022_NE") %>%
    rename(id = multichar_response) %>%
    mutate(cond = factor(title,
                         levels = c("2022_NS", "2022_HJ", "2022_HP"),
                         labels = c("none", "jazz", "pop")
                         )
           )

## Check the results:
dat1 %>% select(-id) %>% head()
summary(dat1)

levels(dat1$cond)
dat1 %$% table(cond, title)
```

Now, we're ready to fit the models.

---

# Modeling

---

Before estimating the hypothesized models, we need to check the influence of the
nesting. So, we'll calculate the ICC to see how much variability is explained by
the individuals' self-similarities.

```{r icc}
## Estimate a random intercept-only model to calculate ICC:
fit0 <- glmer(correct ~ (1|id), data = dat1, family = "binomial")
summary(fit0)
icc(fit0)
```

The clustering doesn't account for very much variability (i.e., `r round(100 * icc(fit0)$ICC_unadjusted, 2)`\%), but we will still control for it via a random
intercept.

```{r model1}
## Fit the hypothesized model:
fit1 <- glmer(correct ~ cond + (1|id), data = dat1, family = "binomial")
summary(fit1)
fixef(fit1) %>% exp()
```

```{r echo = FALSE}
cf <- summary(fit1)$coefficients

b1 <- fixef(fit1)[2] %>% round(3)
z1 <- cf[2, 3] %>% round(2)

b2 <- fixef(fit1)[3] %>% round(3)
z2 <- cf[3, 3] %>% round(2)
```

Pop music produces significantly better results than no music (
$\beta = `r b2`$,
$OR = `r exp(b2) %>% round(2)`$,
$z = `r z2`$,
$p `r getp(fit1, 3)`$
), but results under jazz music and no music do not significantly differ (
$\beta = `r b1`$,
$OR = `r exp(b1) %>% round(2)`$,
$z = `r z1`$,
$p `r getp(fit1, 2)`$
).

```{r model2}
## Change the reference group to evaluate the other comparison:
fit2 <- dat1 %>%
    mutate(cond = relevel(cond, ref = "jazz")) %$%
    glmer(correct ~ cond + (1|id), family = "binomial")
summary(fit2)
fixef(fit2) %>% exp()
```

```{r, echo = FALSE}
cf <- summary(fit2)$coefficients

b2 <- fixef(fit1)[3] %>% round(3)
z2 <- cf[3, 3] %>% round(2)
```

Pop music also produces significantly better results than jazz music  (
$\beta = `r b2`$,
$OR = `r exp(b2) %>% round(2)`$,
$z = `r z2`$,
$p `r getp(fit2, 3)`$
).

---

# Diagnostics

---

Before we can interpret these findings, we need to check out model's assumptions.

```{r assumption1}
## Estimate (deviance) residuals and fitted values (on the latent scale):
res  <- resid(fit2, type = "deviance")
yHat <- predict(fit2, type = "link")

## Plot residuals against fitted values:
data.frame(yHat, res) %>%
  ggplot(aes(yHat, res)) +
  geom_point() +
  geom_smooth() +
  theme_classic()
```

The deviance residuals don't indicate any unmodeled nonlinearity.

```{r assumption2}
## Check homogeneity of residual variances:
resid(fit0, type = "deviance") %>% tapply(dat1$cond, var)

## Check homogeneity and linearity:
data.frame(residual  = resid(fit0, type = "deviance"),
           condition = dat1$cond) %>%
  ggplot(aes(condition, residual)) +
  geom_boxplot() +
  theme_classic()
```

The residual variances appear to be approximately equal in each condition.

---

# Conclusions

---

Overall, we can make the following conclusions.

1. The assumptions seem to be satisfied.
1. The clustering effect is not very strong.
1. Pop music is significantly better than no sound.
1. Pop music is significantly better than jazz music.
1. There is no significant difference between no sound and jazz.

---
